---
title: "Covid Cases in Toronto"
author: "Mark Edney"
date: "07/01/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

## Objective

The objective of this presentation is to present the Covid cases data from [Open
Toronto](https://open.toronto.ca/dataset/covid-19-cases-in-toronto/) using the Plotly
package. 

## Installing Packages

The following code is used to install the necessary packages

```{r packages, results='hide'}
install.packages("opendatatoronto", repos = "http://cran.us.r-project.org",
                 dependencies = TRUE)
library(opendatatoronto)
library(dplyr)
library(plotly)
```

## Load Data

The following code is used to download the data as provided by the Open Toronto 
Database:

```{r data}
	
# get package
package <- show_package("64b54586-6180-4485-83eb-81e8fae3b8fe")

# get all resources for this package
resources <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")
datastore_resources <- filter(resources, tolower(format) %in%
                                      c('csv', 'geojson'))
	
# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
```

## Data cleaning

The following code is used to clean and present the data

```{r clean}
data$`Reported Date` <- as.Date(data$`Reported Date`)
data2 <- data %>% count(`Reported Date`, `Age Group`)
names(data2)[3] <- "Count"
plot1 <- plot_ly(data = data2, y = ~Count, x = ~`Reported Date`,
                 color = ~`Age Group`, type = "scatter")
```

The final plot will be on the next slide

## Plotly Plot

```{r plot, echo=FALSE}
plot1
```
