---
title: "Toronto Bike Rental"
author: "Mark Edney"
date: "14/01/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

## Bike Rental Shiny App

This application use the data collected from the [Toronto Open Data](https://open.toronto.ca/dataset/bike-share-toronto-ridership-data/) to 
generate a histogram of the usage of rental bikes in Toronto during the month of 
June in 2020.
```{r packages, results='hide'}
install.packages("opendatatoronto", 
                 repos = "http://cran.us.r-project.org",
                 dependencies = TRUE)
library(opendatatoronto)
library(tidyverse)
library(lubridate)
library(shiny)
```
## UI

There are two user inputs on the UI side:

- A slider that limits the maximum and minimum of the displayed values
- A checkbox that excludes users with a annual bike pass

```{r UI, eval=FALSE}
        sidebarPanel(
            sliderInput("dur",
                        "Trip Duration:",
                        min = 0,
                        max = 500,
                        value = c(0,500)),
            checkboxInput("freq",
                        "Exclude annual users:",
                        value = FALSE))
```

## Server

```{r Serverhide, echo = FALSE}
 # get package
    package <- show_package("7e876c24-177c-4605-9cef-e50dd74c617f")
    
    # get all resources for this package
    resources <- list_package_resources("7e876c24-177c-4605-9cef-e50dd74c617f")
    # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
    datastore_resources <- filter(resources, tolower(format) %in% c('zip', 'geojson'))
    # load the first datastore resource as a sample
    data <- filter(datastore_resources, name == "Bike share ridership 2020") %>% get_resource()
    data2 <-  data$`2020-06.csv`
    data2[grepl("Time",names(data2))] <- 
        lapply(data2[grepl("Time",names(data2))], parse_date_time, orders = "mdy HM")
    data2$Dur <- as.numeric(data2$End.Time - data2$Start.Time,units="mins")

```

The server cleans up the data and calculates the difference in usage between annual
pass renting versus casual pass renting.

```{r servershow}
    anni <- data2$Dur[data2$User.Type=="Annual Member"]
    cas <- data2$Dur[data2$User.Type=="Casual Member"]
    dif <- mean(anni) -mean(cas)
```

The server also takes the values input from the UI and applies them

```{r server2, eval=FALSE}
        minx <- input$dur[1]
        maxx <- input$dur[2]
        ifelse(input$freq==TRUE,
               data <- data2[data2$User.Type == "Casual Member",],
               data <- data2)
```

## Sample Output

Here is a sample plot of the data and the difference in average usage is : `r round(dif)`
 minutes. Casual riders are more likely to travel for a longer time then frequent
 (annual) riders. 
 
```{r output, echo=FALSE}
        data <- data2
        g <- ggplot(data, aes(x=Dur, fill = User.Type)) + geom_histogram() + 
           xlim(c(0,300)) + labs(x="Trip Duration", y="Count")
g
```

